\documentstyle{article}
\parindent=0pt
\parskip=4pt

\begin{document}

\title{Disk Scheduling in EROS}
\author{
Jonathan S. Shapiro\\
{\tt shap@aurora.cis.upenn.edu}\\
\\
Norman Hardy\\
{\tt norm@netcom.com}}
\date{December 29, 1993}
\bibliographystyle{plain}
\maketitle
\begin{abstract}
The requirements for real-time and reliable system design have some
interesting cross connections.  In this paper we consider the problem
of managing shadowed block I/O in a real time system.  In particular, we
address the issue of maintaining a real-time guarantee in the face of
underlying media deterioration.
\end{abstract}


\section{Introduction}
EROS is a pure capability system closely derived from KeyKos.
\cite{kk:nano,kk:osr}  It presents a secure single-level store with
real-time guarantees to all processes, and provides for failure
restart and reliability.\cite{kk:ckpt}

As part of the reliability design, EROS provides OS support qfor
shadowed disk regions.  Such disk regions are served by two (or more)
independent disk devices, either of which may respond to a read
request for that region.  While shadowing reduces a significant source
of system failures, it significantly complexifies handling of read
faults.

Simultaneously, the EROS system provides access time guarantees for
high priority page I/O.  The convolution of these features imposes
an interesting requirement on the disk subsystem.


\begin{thebibliography}{99}
\bibitem{kk:nano}
Bomberger et. al.
{\em The KeyKOS Nanokernel Architecture},
Proceedings of the USENIX Workshop on Micro-kernels and Other Kernel
Architectures,
USENIX Association,
April, 1992,
pp. 95--112.

\bibitem{kk:osr}
Hardy, Norman,
{\em The KeyKOS Architecture}
Operating Systems Review,
October, 1985,
pp. 8--25.

\bibitem{kk:ckpt}
Landau, Charles R.
{\em The Checkpoint Mechanism in KeyKOS}
Proceedings of the Second International Workshop on Object Orientation
in Operating Systems, 
IEEE,
September, 1992,
pp. 86--91.

\end{thebibliography}

\end{document}




